---
title: "Optimizing Bed Allocation in Stroke Units: A Simulation-Based Analysis"
echo: FALSE
format: pdf
bibliography: reference.bib
csl: academic-medicine.csl
---

### ***Abstract***:

#### **Introduction**:

Geographical stroke units offer effective treatment for ischemic and hemorrhagic stroke patients. Optimal care is achieved when patients are allocated to these wards. Determining the optimal number of beds required is a complex task. This study/project leverages discrete-event simulation, grounded in queueing theory, to ascertain the ideal number of beds for optimal care given patient influx. The study utilized stroke unit data including arrival rates, bed occupation times, and stroke severity probabilities to generate realistic scenarios for the simulation model. The objective of this project was to evaluate the impact of varying bed capacities on the performance of stroke care units such as bed utilization, maximum queue length, proportion of patients waiting, and average queue length.

#### **Method and Design**:

The simulation model considered factors such as patient arrival rates, stroke severity probabilities, and bed occupation times. Patient arrival rates were estimated to be 6 patients/day, based on an annual patient load of \~2000. The distribution of stroke severity was as follows: very mild (probability = 0.2), mild (probability = 0.2), moderate (probability = 0.35), severe (probability = 0.1), and very severe (probability = 0.15). Bed occupation times ranged from 1 day (very mild) to 4-10 days (severe). The study explored 25-50 bed scenarios and considered an increase in patient arrival to 2500 in the near future. Simulation was run for 100 replications, evaluating performance measures (e.g., average queue length, maximum queue length, and capacity utilization) with an initial 25-bed provision.

#### **Discussion:**

The study utilized stroke unit data, including arrival rates, bed occupation times, and stroke severity probabilities, to generate realistic scenarios for the simulation model. A simulation model showed that the initial 25-bed provision was inadequate for the annual patient load of \~2000, leading to prolonged wait times and queues. The percent of patients queuing was around \~ 33% for 25 beds, 23% for 30 beds, and \~5% for 40 beds.

### ***Background and Motivation***

**Introduction to hospital capacity planning**

Healthcare systems worldwide are under continual pressure to provide accessible and effective health services to patients; consequently, the decision-making regarding the adequate provision of services to the patients is crucial for optimal patient care in hospitals. Hence, hospital capacity planning has emerged as a critical issue due to the increasing demand for hospital care and rising costs of inpatient care. Key metrics, such as bed capacity and associated indices such as bed occupancy and the ratio of beds to population are used to determine the availability of inpatient care [@Ravaghi2020]. The number of beds in a hospital can be considered a capital stock, influenced by the performance of medical staff and equipment [@healthGlance2017] [@liu2012] .

A shortage of available beds is the primary cause of delays in emergency admission and surgery cancellations, and premature patient discharge [@rechel2010hospital]. Conversely, underutilization of beds in the hospitals leads to additional hospital costs.

According to Storey et al. (2010) in their article "Hospital capacity planning: From measuring stocks to modelling flows" published in the Medical Journal of Australia, an 85% bed occupancy rate is a reasonable target for hospitals. However, this target may need adjustment depending on the facility and time periods @martin_storey_2010. Effective hospital resource management, including bed capacity, requires short- and long-term hospital capacity planning.

**Stroke cases in Australia and Victoria**

Stroke is ranked as the second leading cause of death with an annual mortality rate of over 6.5 million worldwide and 143.23 million disability-adjusted life years (DALYs) in 2019 according to the *Global stroke fact sheet* released by the [World Stroke Organization](https://www.world-stroke.org/news-and-blog/news/wso-global-stroke-fact-sheet-2022). In Australia in 2018, an estimated 387,000 Australians aged 15 and over (1.3% of the population) had experienced a stroke at *some point* in their lives, based on self-reported data from the [ABS Survey of Disability, Ageing and Carers (ABS 2019)](https://www.aihw.gov.au/reports/heart-stroke-vascular-diseases/hsvd-facts/contents/heart-stroke-and-vascular-disease-subtypes/stroke). As per the same source, in 2020, there were an estimated **39,500 stroke events** in Australia--- more than 100 everyday. The rate of stroke events was 154 per 100,000 population. These stroke events include the new and recurrent stroke.[^1] In 2020, stroke was recorded as the underlying cause of 8,200 deaths, accounting for 5.1% of all deaths in Australia. Stroke was one of the 5 leading causes of death in Australia -- on average, 22 Australians died of stroke each day in 2020.[^2]

[^1]: There are no direct national data sources on the annual number of strokes. However, a related measure can be used as an estimate---the number of stroke events---developed by the AIHW using unlinked hospital and deaths data (AIHW 2022)

[^2]: These figures were obtained from the [Australian Institute of Health & Welfare (AIHW)](https://www.aihw.gov.au/reports/heart-stroke-vascular-diseases/hsvd-facts/contents/summary-of-coronary-heart-disease-and-stroke/stroke)

Since 1980 the death rates for stroke have fallen by three-quarters -- from *110 to 23* per **100,000** population for males, and from *99 to 24* per **100,000** population for females. Access to stroke care units has been attributed to as one of the leading factors for declining death rates in Australia[^3]. Hence adequate access to the stroke care is crucial for the recovery and rehabilitation of the stroke patients as well as in reducing the stroke-related death rates. While, according to the *Victorian Population Health Survey* (2016) [^4], the prevalence of stroke among the adult Victorian population was 2.7% of the adult population.

[^3]: See 'Chapter 4 Changing patterns of mortality in Australia since 1900' in [*Australia's health 2022: data insight*](https://www.aihw.gov.au/reports/australias-health/australias-health-2022-data-insights).

[^4]: 2.7% of the adult Victorian Population self-reported doctor diagnosed stroke [in the survey](https://www.health.vic.gov.au/population-health-systems/victorian-population-health-survey-2016).

In Melbourne, the incidence and prevalence of stroke follow similar trends observed across Australia. The growing population and aging demographic contribute to the increasing demand for stroke care in the region. According to the Australian Institute of Health and Welfare, the age-standardized prevalence of stroke in Melbourne has been on the rise, highlighting the need for better access to specialized care for stroke patients.

#### Monash Medical Research Institute

In the stroke ward at the Monash Hospital there are a total of 46 beds as used by the neurology, neurosurgery and stroke patients. The objective of this project was to determine the optimal number of beds specifically for the stroke patients at the Monash Hospital to ensure the optimum care for the stroke patients.

For the purpose of this project at the *Monash Medical Research institute*, I have used the **simulation modelling** to represent a queuing system within a hospital setting. The main objective of this simulation was to simulate the flow of patients as they arrive, wait in line (queue) for an available bed, receive treatment, and exit the resource (making the bed available for the next patient in the queue to use).

### *Objectives and Significance*

#### **Objectives**

The primary objective of this project was to develop *discrete event simulation* (DES) model that accurately captures the dynamics of the stroke ward at the Monash hospital ensuring that complex interactions between patients and beds are analyzed and provides an optimal number of beds needed at the stroke ward to ensure that the stroke patients receive optimum care. In order to achieve these objectives an evaluation of different bed capacities on the proportion of the patients waiting to be admitted into the ward and the utilization rates were considered. One of the goals was also to provide decision support tool to the hospital administrators and policymakers in optimizing bed allocation at the stroke ward.

#### **Significance and Contributions**:

In this project, *Discrete Event Simulation* has been employed to develop a queuing system that accurately captures the intricate complexity and inherent variability of a stroke ward. This model takes into account the stochastic nature of the patient arrivals, stroke severity levels, and length of stay, enabling a comprehensive analysis of the hospital ward's performance under different scenarios. To ensure the model's accuracy I incorporated key parameters such as the arrival rates, bed occupation times and the stroke level probabilities which were acquired from a combination of historical data, hospital statistics and relevant literature. @green2011queueing By effectively representing the complex interaction between these factors, this DES model is meant to provide valuable insights into the stroke ward's operations at the Monash Hospital. Furthermore, this model enables users to generate performance metrics, such as the percent of patients waiting at the given number of beds in the unit for a specific patient population, or the utilization capacity for a particular at a given number of beds for the selected patient load in the care unit under variety of scenarios. This simulation model is designed to be used as a support tool for hospital administrators when determining optimal bed allocation and resource management, ultimately enhancing patient care and outcomes.

#### **Novel Features:**

While simulation modeling in healthcare is not a novel concept, this project aims to broaden the accessibility of the simulation modelling developed as part of this project by making it available as a support tool to a wider audience through a user-friendly shiny web-app which can be accessed [here.](https://aryan-sultan-1120.shinyapps.io/App_for_stroke_simulation/) (https://aryan-sultan-1120.shinyapps.io/App_for_stroke_simulation/). The Shiny app enables users to interactively adjust the number of patients and beds using slider inputs, allowing them to visualize a plot of the percentage of patients waiting based on the selected number of patients and beds as well as a plot of the utilization rates at the selected number of beds. Furthermore, the users are able to see the optimal number of beds for the selected patient load. By offering this tool in a web-based format, a wider audience can benefit from the insights generated by the simulation model, thereby enhancing decision-making processes in healthcare settings.

### *Methodology*

The methodology employed in this analysis is based on Discrete Event Simulation (DES), a widely recognized and effective technique for modeling complex systems with stochastic elements. DES was chosen for this project due to its ability to accurately represent the intricacies of a stroke care unit, incorporating the variability and uncertainty associated with patient arrivals, stroke severity levels, and length of stay.

#### **Discrete Event Simulation (DES)**

Discrete Event Simulation (DES) is a computer-based modeling technique that simulates the operation of a system as a sequence of discrete events in time. Each event represents a change in the system's state.

DES offers several advantages over traditional analytic methods in that it can model a variety of systems, including those with the stochastic and dynamic elements that may otherwise prove difficult to be analyzed using traditional methods such as a deterministic queuing models, moreover, DES allows for representing the individual components of the system and the intricate interaction between these components in greater detail.

The stroke ward simulation model, for example, incorporates various stochastic and dynamic elements, such as patient arrivals, stroke severity levels, and length of stay. These elements are inherently unpredictable and can change over time. Traditional analytic methods, such as deterministic queuing models, may not fully capture the randomness and variation in these components. DES, on the other hand, is specifically designed to model such types of the systems. By using random number generation to simulate the variability in patient arrivals, stroke severity, and length of stay, DES allows for a creation of a realistic model, which allows for a more reliable assessment of the ward's performance under different scenarios, such as changes in patient arrival rates or bed capacities.

Similarly, the simulation model in this study consists of multiple components, including patients, beds, and resources. Each of these components interacts with one another in a complex way, as patients move through the ward, occupy beds, and utilize resources. Traditional analytic methods might simplify these interactions or aggregate them, potentially losing critical information about the system's behavior. DES, however, represents individual components and their interactions. For example, it can track each patient as they enter the ward, get assigned to a bed, receive treatment, and eventually get discharged. This level of detail allows the simulation to provide a more accurate representation of the system's behavior.

Addressing the stochastic and dynamic elements and accurately representing individual components and interactions of the system are important for generating more reliable results. Furthermore, one of the key advantages of the DES is the flexibility. DES model allows for changes in the system's behavior to be successfully incorporated into the model, and even test for various scenarios and interventions. For example in my simulation model, the ***changes*** in patient population (inflow), their length of stay and stroke level probabilities associated with the stroke categories can be accounted for easily and reliably.

In the case of the stroke ward simulation, DES was a better choice because it allowed for the modeling of complex interactions between patients, beds, and resources in a dynamic and stochastic environment. The simulation was able to capture patient arrivals, bed occupancy, and other factors, providing a more accurate representation of the system's behavior and helping to identify optimal resource allocation. @jacobson2006discrete

#### Data

The data used in this project is derived from multiple sources including historical records and hospital statistics, which included arrival rates, bed occupation times, stroke categories and associated probabilities. Furthermore, valuable insights regarding the patient arrivals, annual patient load and the bed occupation times for different categories of stroke patients were obtained from the supervisor for this project, Professor Thanh Phan, Head of Neuroscience Research, Monash Health.

It is important to note that the data utilized in this project may have certain limitations, such as potential inaccuracies or biases in the original records. However, these limitations have been carefully considered and addressed wherever possible to ensure the reliability of the analysis. Furthermore, consultations with and feedback from supervisor for this project, Professor Phan were acquired regularly to ensure the reliability of the results generated from this model as well as the accuracy of the information and data used in this project. Moreover, the steps were taken to mitigate any potential effects of inaccuracies that emerge as a result of using simulation model to reflect the real-life scenarios of the stroke ward. It is important to note that simulation models presents certain limitations.

One of the key limitations of the simulation model is the validation and calibration of simulation model is a challenging process and can be quite time consuming at times, particularly in the absence of stroke specific data. The information used in this simulation model was collected from multiple wards where the stroke patients were admitted; there were a total of ten wards that had the stroke patients; the total number of these patients and their bed occupation times for the time period: October 2022 and March 2023 were observed and used to validate the length of stay of the patients from the simulation model. Furthermore, the critical information on patient's length of stay (bed occupation times), stroke category probabilities were supplied by the supervisor for this project at Monash Health, Professor Thanh Phan. These information on the stroke patients were incorporated into the model.

One of the other challenges I faced was that the computation times were long. DES can be computationally intensive when running multiple thousand simulation or multiple scenarios. This drawback of the DES made it difficult to run multiple simulations and compute performance measures of each simulation and average those results to produce more reliable results for the performance metrics.

#### Method

In order to develop this model several functions were created to generate the patient inter-arrival times, assign stroke categories to patients in the simulation, generate their length of stay based on the stroke category they were assigned, simulate the patient inflow, stay and exit, and finally run multiple simulations and for each simulation compute certain performance metrics such as the percent of the patients waiting, utilization rates and maximum queue length. The stroke categories were assigned based on the stroke level probabilities. The total annual patient load considered for this analysis was \~ 2000. The performance metrics were computed for 2000 patients as well as for the future estimated number of 2500 patients.

The following parameters were used to lay the foundation of the stroke simulation model. These parameters were obtained from the historical data of the stroke ward.

-   Total annual patients: 2000

-   Arrival rate: $\frac {2000} {365} = 6$ patients per day

-   Stroke level probabilities & bed occupation times see Table 1:

```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

#| label: tbl-strokeCat
#| tbl-cap: Stroke Categories, Probabiliy and Bed Occupation times

library(tidyverse)
library(ggplot2)
library(bookdown)

set.seed(2023)
# Parameters

total_patients <- 2000 # assuming there are 2000 all around the year
total_patients_future <- 2500 # assuming the 2500 patients in the future 
patients_per_month <- 2000/12 # patients per month
arrival_rate <- total_patients/365  # average 6 patients per day
arrival_rate_future <- total_patients_future/365 # average 7 patients per day

# Bed occupation time (in days) for each stroke level

bed_occupation_time <- list(c(4.43), c(4.46), c(6.67), c(6.92), c(7.74), c(9.29))


# Probability of each stroke level
stroke_level_prob <- c(0.0570,0.342, 0.217, 0.136, 0.0971, 0.151)

# visualization

library(ggplot2)

bed_occupation_time_df <- data.frame(
  stroke_category = c("Very Mild", "Mild", "Mild-Moderate", "Moderate", "Moderate-Severe", "Severe"),
  bed_occupation_time = c(4.43,4.46,6.67,6.92,7.74,9.29),
  probability = c(0.0570,0.342, 0.217, 0.136, 0.0971, 0.151)
)

updated_bed_occupation_time_df <- bed_occupation_time_df

colnames(updated_bed_occupation_time_df) <- c("Stroke Categories", "Bed Occupation times", "Stroke Probability")

updated_bed_occupation_time_df %>%
  knitr::kable(caption = "Stroke categories and corresponding length of stay and probabilities")
```

See Figure:1 for the summary of the stroke categories and the associated probabilities with corresponding length of stay.

```{r figcat, fig.cap="Stroke categories, probabilities and bed occupation times", fig.align= 'center', fig.height=4, fig.width= 6, eval=TRUE}

# viz for the stroke categrories.

v1 <-ggplot(updated_bed_occupation_time_df, aes(x = `Stroke Categories`, y = `Bed Occupation times`)) +
  geom_col(fill = "steelblue") +
  theme_minimal()+
  geom_text(aes(label = scales::percent(`Stroke Probability`, accuracy = 0.1)), 
            position = position_stack(vjust = 0.5), color = "white", size = 4) +
  theme_minimal() +
  labs(
       x = "Stroke Category",
       y = "Bed Occupation Time (days)")
v1

```

With these input parameters, the stroke simulation model was developed to achieve the results by creating functions that generated the inter-arrival times, assigned stroke categories to the patients in the simulation, and generated the bed occupation times for the patients based on their stroke categories.

#### Inter-arrival times

In specifying a queuing model, we make assumptions about the probabilistic nature of the arrival and service processes. The most common assumption to make about arrivals is that they follow a *Poisson process. @green2011queueing*

Assume $N(t)$ is the number of arrivals during a time period of duration $t$ and $N(t)$ has a Poisson distribution, then:

Probability= $$\{N(t) = n\} = \frac {e-λt
 (λt)n} {n!} $$ where $\lambda$ is called the *rate* and is the expected number of arrivals per unit time. Another way to characterize the Poisson process is that the time between consecutive arrivals, called the *inter-arrival time*, has an exponential distribution.

Mathematically, the inter-arrival time for our model was generated in the following way:

-   Total number of patients = 2000, hence let $n=2000$

-   Arrival rate = 6 patients, therefore, $\lambda =6$

    The exponential distribution has a probability density function (PDF) given by:

    $$f(T) = λ \times Exp(-λ(T))$$

    where $\lambda$ is the arrival_rate, and $T$ is a random variable representing the time between two successive patient arrivals.

    In the simulation for each patient "$i$" in the 2000 patients (total number of patients), we generate an inter-arrival time $T_i$ from an exponential distribution with the rate $(\lambda)$:

    $$T_i \sim Exp(6)$$

    This generates a sequence of inter-arrival times: $T_1, T_2, T_3,… T_{2000}$ that is exponentially distributed with $\lambda$ = 6.

    In R, this implementation is shown below:

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

# total patients = 2000
# arrival rate = 6

generate_interarrival_times <- function(total_patients, arrival_rate) {
  interarrival_times <- rexp(total_patients, rate = arrival_rate)
  return(interarrival_times)
}

```

#### Assigning the stroke categories to Patients

Next, I created a `assigning_stroke_levels` function to assigns stroke categories to patients.

The `assign_stroke_levels` function assigns stroke categories to patients based on a given set of probabilities for each category. These probabilities have been mentioned earlier in the report, please refer to Table 1. The function takes two arguments:

1\. `total_patients`: The total number of patients for which stroke categories will be assigned.

2\. `stroke_level_prob`: A vector containing the probability of each stroke category.

Mathematically, this process can be explained as follows:

Let $Y_1, Y_2, ..., Y_n$ be the stroke categories for $n$ patients, where $n$ =2000. Each $Y_i$ is an independent random variable taking values from $1 \to k$, where $k=6$ is the number of stroke categories . The probability of $Y_i$ taking the value $j$ : $(1 ≤ j ≤ 6)$ is given by:

$P(Y_i = j) = p_j$

where $p_j$ is the probability of the $j^{th}$ stroke category, specified in the `stroke_level_prob` vector.

Then in this R-function stroke categories are assigned to patients by sampling from a discrete probability distribution with the specified probabilities:

$Y_1, Y_2, ..., Y_n$ = sample( $1:k$ , size = $n$ , replace = TRUE, prob = $p$)

The `replace = TRUE` argument ensures that each patient's stroke category is assigned independently, allowing for repeated selections of the same category.

The implementation in R is as follows:

```{r eval=TRUE, echo=TRUE}

# stroke_level_prob <- c(0.0570,0.342, 0.217, 0.136, 0.0971, 0.151)

assign_stroke_levels <- function(total_patients, stroke_level_prob) {
  stroke_levels <- sample(1:length(stroke_level_prob), 
                          size = total_patients, 
                          replace = TRUE, 
                          prob = stroke_level_prob)
  return(stroke_levels)
}

stroke_levels <- assign_stroke_levels(total_patients, stroke_level_prob)
```

#### Generating the Length of Stay

As discussed (refer to Table 1), the bed occupation times for each category are defined. For this simulation model once the inter-arrival times and the stroke categories to the arriving patients have been assigned, the next step was to generate the length of stay of these patients based on their stroke level/category.

In the context of our stroke simulation model, the length of stay is computed using a piecewise constant function that takes the stroke category assigned to each patient as input and maps it to the corresponding bed occupation time. The function can be described mathematically as follows:

Let $f(Y)$ be the piecewise constant function that represents the length of stay for a patient with stroke category $Y$. The stroke categories are indexed by $i$, with each category corresponding to a bed occupation time $c_i$. The function $f(Y)$ is defined as:

$f(Y) = c_i$, if $Y$ belongs to the $i^{th}$ stroke category.

In the R function `generate_length_of_stay`, the length of stay for each patient is implemented using this piecewise constant function. The function iterates through all patients, assigning a length of stay based on their stroke category:

```{r eval=TRUE, echo=FALSE}
generate_length_of_stay <- function(total_patients, 
                                    bed_occupation_time, 
                                    stroke_levels) {
  length_of_stay <- numeric(total_patients)
  
  for (i in 1:total_patients) {
    
    stroke_level <- stroke_levels[i]
    
    if (length(bed_occupation_time[[stroke_level]]) == 1) {
      
      length_of_stay[i] <- bed_occupation_time[[stroke_level]][1]
    } 
    else {
      length_of_stay[i] <- sample(bed_occupation_time[[stroke_level]][1]
                                  :bed_occupation_time[[stroke_level]][2], 1)
    }
  }
  
  return(length_of_stay)
}


# Generating length_of_stay for each patient
length_of_stay <- generate_length_of_stay(total_patients, 
                                          bed_occupation_time, 
                                          stroke_levels)
```

1\. First this function, retrieve the stroke category for the patient, denoted by `stroke_level`.

2\. Then it checks if the bed occupation time for that category is a single value or a range of values.

3\. If it's a single value, this function assigns the length of stay to that value: `length_of_stay[i] <- bed_occupation_time[[stroke_level]][1]`.

4\. If it's a range of values, however, then it randomly sample a value from that range and assigns it as the length of stay: `length_of_stay[i] <- sample(bed_occupation_time[[stroke_level]]`

`[1]:bed_occupation_time[[stroke_level]][2], 1)`.

This way, the R function `generate_length_of_stay` computes the length of stay for each patient based on their stroke category using the piecewise constant function described above.

#### Arrivals data

Having defined these functions, the next step is check the distribution of the length of stay of patients in our simulation model and have a look at the patients arrival data.

```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
set.seed(2023)

# generating the arrival data for the current volume of patients

# Add number_of_patients and length_of_stay to arrivals_data

interarrival_times <- generate_interarrival_times(total_patients, arrival_rate)

stroke_levels <- assign_stroke_levels(total_patients, stroke_level_prob)

length_of_stay<- generate_length_of_stay(total_patients, bed_occupation_time, stroke_levels)

arrivals_data <- data.frame(interarrival_times, stroke_levels, length_of_stay)

```

```{r eval=TRUE, echo=TRUE, warning=FALSE}
arrivals_data$stroke_categories <- factor(arrivals_data$stroke_levels,
                                          levels = c(1, 2, 3, 4, 5, 6),
                                          labels = c("Very Mild", "Mild", "Mild-Moderate", "Moderate", "Moderate-Severe", "Severe"))

colnames(arrivals_data) <- c("Interarrival times", "stroke levels", "Length of stay", "stroke categoryNames")

arrivals_data %>% head() %>%
  knitr::kable(caption = "Patient Arrival Data")
```

This table (Table: 2) shows us just the first six rows of the actual data I created with the three functions mentioned earlier.

The distribution of the bed occupation times (LOS) of the patients in our simulation can be viewed in Figure 2.

```{r figtwo, fig.cap="Patients Arrival Data", fig.align= 'center', fig.height=4, fig.width= 6, eval=TRUE}
ggplot(arrivals_data, aes(x = `stroke categoryNames`, y = `Length of stay`, fill = `stroke categoryNames`)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  theme_minimal() +
  scale_y_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10))+
  labs(title = "Mean Length of Stay by Stroke Category",
       x = "Stroke Category",
       y = "Mean Length of Stay (days)") +
  scale_fill_discrete(name = "Stroke Category")+
        theme(axis.text.x = element_text(angle = 45 ,hjust = 1),
            legend.position = "none")
```

In Figure: 2 one can see the mean length of the patients' stay in the stoke ward against the stroke categories/levels.

#### Stroke Simulation

### References

### 
